---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# socratadata

<!-- badges: start -->
<!-- badges: end -->

Explore Socrata data with ease.

`socratadata` provides a modern interface for downloading data from [Socrata](https://socrata.com) open data portals powered by Rust. `socratadata` improves upon the existing [`RSocrata`](https://dev.socrata.com/connectors/rsocrata) package by introducing support for all [Socrata datatypes](https://dev.socrata.com/docs/datatypes/), as such `socratadata` natively supports datasets with geospatial fields by returning them as `sf` objects when applicable.

## Installation

You can install the development version of socratadata from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("ryanzomorrodi/socratadata")
```

## Example

Use `list_socrata()` to explore the datasets available on any socrata data portal.

```{r}
library(socratadata)

catalog <- list_socrata("https://data.cityofchicago.org")
print(catalog)
```

Use `read_socrata()` to read a socrata dataset into R.
```{r}
cta_ridership <- read_socrata(
  "https://data.cityofchicago.org/Transportation/CTA-Ridership-Daily-Boarding-Totals/6iiy-9s97/about_data"
)
print(cta_ridership)
```

Spatial data will be read as an `sf` object.
```{r}
chi_community_areas <- read_socrata(
  "https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Community-Areas/igwz-8jzy/about_data"
)
print(chi_community_areas)
```
