---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# socratadata

<!-- badges: start -->
<!-- badges: end -->

Explore Socrata data with ease.

`socratadata` provides an easy-to-use interface for downloading data from [Socrata](https://dev.socrata.com/) open data portals powered by Rust. `socratadata` improves upon the existing [`RSocrata`](https://dev.socrata.com/connectors/rsocrata) package by introducing support for the [Socrata Discovery API](https://dev.socrata.com/docs/other/discovery#?route=overview) and all [Socrata datatypes](https://dev.socrata.com/docs/datatypes/).

Unlike `RSocrata`, `socratadata` does not support uploading or editing existing datasets.

## Installation

You can install `socratadata` from CRAN.

```r
install.packages("socratadata")
```

You can install the development version of socratadata from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("ryanzomorrodi/socratadata")
```

## Usage

### Search for datasets

Use `soc_discover()` to explore the datasets with a domain.

```{r setup}
library(socratadata)

soc_discover(domains = "data.cityofchicago.org")
```

Or with a query.

```{r}
soc_discover(query = "bus")
```

Or with categories.

```{r}
soc_discover(categories = "transportation")

soc_discover(
  domains = "data.cityofchicago.org",
  domain_category = "Transportation"
)
```

Or with tags.

```{r}
soc_discover(
  domains = "data.cityofchicago.org",
  domain_tags = "public transit"
)
```

Or with ids.
```{r}
soc_discover(ids = c("6iiy-9s97", "pnau-cf66"))
```

### Download data

`socratadata` supports unauthenticated requests via the v2.1 API. It is, however, recommended that you [obtain an api key](https://support.socrata.com/hc/en-us/articles/210138558-Generating-App-Tokens-and-API-Keys)
to make your code more future-proof. `soc_read()` will automatically authenticate with the `"soc_api_key_id"` and `"soc_api_key_secret"` environment variables.

You can create an `.Renviron` file and add your keys to it like so:

```txt
soc_api_key_id="your_id_here"
soc_api_key_secret="your_secret_here"
```

And retrieve those keys using 

```r
Sys.getenv("soc_api_key_id")
#> [1] "your_id_here"
Sys.getenv("soc_api_key_secret")
#> [1] "your_secret_here"
```

Use `soc_read()` to read a socrata dataset into R.
```{r}
soc_read(
  "https://data.cityofchicago.org/Transportation/CTA-Ridership-Daily-Boarding-Totals/6iiy-9s97/about_data"
)
```

Spatial data will be read as an `sf` object.
```{r}
soc_read(
  "https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Community-Areas/igwz-8jzy/about_data"
)
```

You can even perform complex queries using [Socrata Query Language (SoQL)](https://dev.socrata.com/docs/queries/) via `soc_query()`.
```{r, message = FALSE}
lower_west_side <- soc_read(
  "https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Community-Areas/igwz-8jzy/about_data",
  query = soc_query(
    where = "community LIKE 'LOWER WEST SIDE'"
  )
)
print(lower_west_side)

cta_ridership <- soc_read(
  "https://data.cityofchicago.org/Transportation/Taxi-Trips-2013-2023-/wrvz-psew/about_data",
  query = soc_query(
    select = "pickup_community_area, count(*) as n",
    where = glue::glue(
      "within_polygon(dropoff_centroid_location, '{sf::st_as_text(lower_west_side$the_geom)}')"
    ),
    group_by = "pickup_community_area",
    order_by = "n DESC"
  )
)
print(cta_ridership)
```

### Extract metadata

Access a dataset's metadata using `soc_metadata()`.
```{r}
soc_metadata(cta_ridership)
```

Or explore a dataset's metadata using it's url.
```{r}
soc_metadata(
  "https://data.cityofchicago.org/Transportation/CTA-Ridership-Daily-Boarding-Totals/6iiy-9s97/about_data"
)
```
