---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# socratadata

<!-- badges: start -->
<!-- badges: end -->

Explore Socrata data with ease.

`socratadata` provides a modern interface for downloading data from [Socrata](https://socrata.com) open data portals powered by Rust. `socratadata` improves upon the existing [`RSocrata`](https://dev.socrata.com/connectors/rsocrata) package by introducing support for all [Socrata datatypes](https://dev.socrata.com/docs/datatypes/) and queries via the [Socrata Query Language (SoQL)](https://dev.socrata.com/docs/queries/).

## Installation

You can install the development version of socratadata from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("ryanzomorrodi/socratadata")
```

## Example

Use `soc_list()` to explore the datasets available on any socrata data portal.

```{r}
library(socratadata)

catalog <- soc_list("https://data.cityofchicago.org")
print(catalog)
```

Use `soc_read()` to read a socrata dataset into R.
```{r}
cta_ridership <- soc_read(
  "https://data.cityofchicago.org/Transportation/CTA-Ridership-Daily-Boarding-Totals/6iiy-9s97/about_data"
)
print(cta_ridership)
```

Spatial data will be read as an `sf` object.
```{r}
chi_community_areas <- soc_read(
  "https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Community-Areas/igwz-8jzy/about_data"
)
print(chi_community_areas)
```

You can even perform complex queries using [Socrata Query Language (SoQL)](https://dev.socrata.com/docs/queries/).
```{r}
lower_west_side <- soc_read(
  "https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Community-Areas/igwz-8jzy/about_data",
  query = soc_query(
    where = "community LIKE 'LOWER WEST SIDE'"
  )
)
print(lower_west_side)

trips_to_lws_by_ca <- soc_read(
  "https://data.cityofchicago.org/Transportation/Taxi-Trips-2013-2023-/wrvz-psew/about_data",
  query = soc_query(
    select = "pickup_community_area, count(*) as n",
    where = glue::glue(
      "within_polygon(dropoff_centroid_location, '{sf::st_as_text(lower_west_side$the_geom)}')"
    ),
    group_by = "pickup_community_area",
    order_by = "n DESC"
  )
)
print(trips_to_lws_by_ca)
```
